# ggplot2 cheatsheet

```{r}
opts_chunk$set(warning=FALSE, message=FALSE)
```

## Loading ggplot

```{r}
library(ggplot2)
```

## Loading sample data

```{r}
set.seed(1410)  # make the sample reproducible
head(diamonds)
dsmall <- diamonds[sample(nrow(diamonds), 100), ]
```

## Basic use (qplot)

```{r}
qplot(carat, price, data=diamonds)
```

## Color, size, shape and other aesthetic attributes

```{r}
qplot(carat, price, data=dsmall, color=color, shape=cut, alpha=I(1/2))
```

## Plot geoms

```{r}
qplot(carat, price, data=dsmall, geom=c("point", "smooth"))
```

### Adding a smoother
There are many different smoother that can be used with `method` argument.

```{r}
qplot(carat, price, data=dsmall, geom=c("point", "smooth"), method="lm")
```

### Boxplots and jittered points

```{r}
qplot(color, price/carat, data=diamonds, geom="jitter")
```

```{r}
qplot(color, price/carat, data=diamonds, geom="boxplot")
```

### Histogram and density plots

```{r}
qplot(carat, data=diamonds, geom="histogram", fill=color)
```

```{r}
qplot(carat, data=diamonds, geom="density", color=color)
```

Change the amount of smoothing with `binwidth` argument.

```{r}
qplot(carat, data=diamonds, geom="histogram", binwidth=1)
qplot(carat, data=diamonds, geom="histogram", binwidth=0.1)
qplot(carat, data=diamonds, geom="histogram", binwidth=0.01)
```

### Bar charts

```{r}
qplot(color, data=diamonds, geom="bar")
```

```{r}
# bar plot of diamond color weighted by carat
qplot(color, data=diamonds, geom="bar", weight=carat) +
    scale_y_continuous("carat")
```

### Time series

```{r}
head(economics)
```

```{r}
qplot(date, unemploy/pop, data=economics, geom="line")
```

### Faceting

```{r}
qplot(carat, data=diamonds, facets=color~.,
      geom="histogram", binwidth=0.1, xlim=c(0,3))
```

### Other options
- `xlim` and `ylim`: set limits for x- and y-axis (e.g. `xlim=c(0,20)`)
- `main`: main title for the plot
- `xlab` and `ylab`: labels for x- and y-axis

```{r}
qplot(carat, price, data=dsmall,
      xlab="Price ($)",
      ylab="Weight (carats)",
      main="Price-weight relationship")
```
